\bf Visual Inference for Significance and Goodness-of-Fit Testing for a Social Network Model

### 2017-10-30 ASApaper/paper-child.Rnw

To determine the threshold at which the effect becomes noticeable, we examined six different levels of the effect, three negative and three positive ones. We hypothesized that the perception of the effect by our participants would look something like Figure~\ref{fig:hypothesis}. The higher in absolute value the parameter is, the more likely participants are to choose the alternative model out of the lineup. To determine the six levels of the effects we want to test, started with the estimates of the parameter at hand as determined by the procedure described in Section~\ref{sec:models}, and multiply it or changing its sign to determine what \textit{we} notice about the effect of the parameter in simulations from the changed models.

In order to determine which specific values of the parameters we wanted to examine, we constructed an online application that created the lineup protocol for us to be the guinea pigs of our own experiment \citep{theqs}. A screen shot of the app wee created with the \texttt{shiny} package is shown in Figure~\ref{fig:shiny} \citep{shiny}. On the left side of the screen, the user\footnote{Please visit \url{https://sctyner.shinyapps.io/saom_lineup_creation/} to create lineups constructed from the models we present for this data for yourself.} can input the information necessary for creating a lineup of the models M1 through M6 for the data in Section~\ref{sec:senatedata}: first, choose to simulate only one plot from the specified model (analogous to changing the alternative model in the lineup protocol) or to simulated $M-1$ plots from the specified model (analogous to changing the null model in the lineup protocol); the model of interest; the wave of the data to examine; if model M1 is selected, whether to alter the density or the reciprocity parameter; the size of the lineup; the amount by which to multiply the effect selected; a random seed for replicability; and a layout algorithm to use for the node-link diagrams. There is also a checkbox if the user wishes the nodes to be colored by the size of the connected component to which they belong. The plots that appear that are \textit{not} from the model specified using the other options are simulations from model M1 with the estimates of the rate parameters, $\beta_1$ and $\beta_2$ given in Table~\ref{tab:fittedvalues}.

Using the ``Picking Lineups" web application, we settled on six parameter values to test for each of our six effects, $\beta_1, \dots \beta_6$. The complete details of the parameters tested using the lineup protocol is given in Table~\ref{tab:experdetail}. In the case of both $\beta_4$ and $\beta_6$, we could not determine a value for negative effects that made the data simulated from M3 and M5 look different than null model simulations from model M1. Because we could not detect these effects at these values, we decided that the less experienced participants in our experiment would also not be able to. So instead of testing the negative values of these effects, we examined a different scenario: we placed 5 simulations from positive values of the parameter with one simulation from model M1 in a lineup. We refer to this later on as the ``reverse" scenario. We used the reverse scenario to determine if the perception of the effect size would be symmetric: if an effect is noticed $x$\% of the time at value $\beta_k = \beta_{k_0}$ when one simulation from the corresponding model is placed among five null plots from model M1, then when five simulations from the model with $\beta_k = \beta_{k_0}$ are put in a lineup with one simulation from model M1, the plot from the simpler model should be noticed about $x$\% of the time as well.

>>>>

To determine the threshold at which an effect becomes noticeable, we examine six different levels of the effect, three negative and three positive ones. Figure~\ref{fig:hypothesis} shows a sketch of what the detection probability by participants' looks like hypothetically with varying effect size: the higher in absolute value the parameter is, the more likely participants are to choose the alternative model out of the lineup. To determine the exact values of the six levels we want to test for each effect, we started with the estimates of the parameter at hand (see  Section~\ref{sec:models}), and used small negative and positive factors to determine at what point \textit{we} noticed the effect of the parameter in simulations from the changed models.

For this we constructed an online application that created the lineup protocol for us to be the guinea pigs of our own experiment \citep{theqs}. A screen shot of the app we created with the \texttt{shiny} package is shown in Figure~\ref{fig:shiny} \citep{shiny}. On the left side of the screen, the user\footnote{Please visit \url{https://sctyner.shinyapps.io/saom_lineup_creation/} to create lineups constructed from the models we present for this data for yourself.} can input the information necessary for creating a lineup of the models M1 through M6 for the data in Section~\ref{sec:senatedata}: first, choose to simulate only one plot from the specified model (analogous to changing the alternative model in the lineup protocol) or to simulated $M-1$ plots from the specified model (analogous to changing the null model in the lineup protocol); the model of interest; the wave of the data to examine; if model M1 is selected, whether to alter the density or the reciprocity parameter; the size of the lineup; the amount by which to multiply the effect selected; a random seed for replicability; and a layout algorithm to use for the node-link diagrams. There is also a checkbox if the user wishes the nodes to be colored by the size of the connected component to which they belong. The plots that appear that are \textit{not} from the model specified using the other options are simulations from model M1 with the estimates of the rate parameters, $\beta_1$ and $\beta_2$ given in Table~\ref{tab:fittedvalues}.

Using the ``Picking Lineups" web application, we settled on six parameter values to test for each of our six effects, $\beta_1, \dots \beta_6$. The complete details of the parameters tested using the lineup protocol is given in Table~\ref{tab:experdetail}. In the case of both $\beta_4$ and $\beta_6$, we could not determine any values for negative effects that made the data simulated from M3 and M5 look different than null model simulations from model M1. Therefore we decided that the lesser experienced participants in our experiment would also not be able to. Instead of testing the negative values of these effects, we are examining a different scenario: we placed 5 simulations from positive values of the parameter with one simulation from model M1 in a lineup. We refer to this later on as the ``reverse" lineup scenario. We used the reverse scenario to determine if the perception of the effect size is symmetric: if an effect is noticed $x$\% of the time at value $\beta_k = \beta_{k_0}$ when one simulation from the corresponding model is placed among five null plots from model M1, then when five simulations from the model with $\beta_k = \beta_{k_0}$ are put in a lineup with one simulation from model M1, the plot from the simpler model should be noticed about $x$\% of the time as well.
